
---
title: "Testing BMAs"
author: "Tom August, Stephen Freeman, Nick Isaac"
date: "14 June 2018" 
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

This version needs to have the data in a folder underneath the working directory
It downloads the latest version of BRCindicators from my repository
It reads the butterfly data from the W drive
modified from my script "butterflies_bms_180608.R""


```{r setup, include = FALSE}
# check you have the latest version
install_package <- FALSE
if(install_package){
  library(devtools)
  install_github(repo = 'drnickisaac/BRCindicators')
}

library(BRCindicators)
library(reshape2)
library(RColorBrewer)
library(ggplot2)

iter <- 250e3
save_model <- TRUE
test <- FALSE

# create a folder for the results
#results_folder <- 'BMA_development/plots'
results_folder <- '../models'
if(!dir.exists(results_folder)) dir.create(results_folder)

#knots <- c(8, 12, 16)
#sf <- c(1,3,9)
```

## Introduction

We are testing the new BMA indicators developed by Steve and Nick and implemented in BRCindicators by Tom.


## Setup
```{r setup dataset, results = 'hide'}
## Load Butterfly data

#Use the butterfly data to compare the methods outputs. We used this butterfly data to compare methods including MSI previously so this is a good dataset to test on. 

# The 2016 butterfly data comes from here:
#dir <- 'W:/PYWELL_SHARED/Pywell Projects/BRC/Tom August/R Packages/BRCindicators/Other_files'
#load(file.path(dir, 'butterfly_data.rdata'))

# the 2017 data is stored locally
ram_data <- read.csv('../data/all_spp_ci.csv')
all_species <- read.csv(file='../data/all_species.csv')

head(ram_data)
head(all_species)

in_data <- ram_data[,c('SPECIES','YEAR','TR0OBS','TR0SE')] 
names(in_data) <- c('species','year','index','se')
nrow(in_data)

in_data$se[in_data$se == 0] <- NA 
# SEs are on the natural log scale so change them to the log10 scale
#in_data$se <- in_data$se/log(10) NO - they are acutally on the correct scale

# rescale year
firstyr <- min(in_data$year)
length(unique(in_data$species))
# start the years at 1
in_data$year <- in_data$year - firstyr + 1
```

  
## Run
Now we can run the models.

```{r runButterflies, echo = FALSE, message = FALSE, fig.width = 8, fig.height = 6, warning = FALSE, results = 'hide'}

#for(group in c('wider')){
#for(group in c('specialist')){
for(group in c('wider','specialist')){
    for(nk in c(8, 10, 12)){
#    for(nk in c(8)){
#        for(scaling_factor in  c(5)){
        for(scaling_factor in  c(1,3,5)){
              in_data$se <- in_data$se * scaling_factor

    m_data <- subset(in_data, species %in% all_species[all_species$type==group, 'BMScode'])

message <- paste0(': ', nk,' knots; se scaling=', scaling_factor)

bma_indicator <- bma(m_data, model = 'smooth_det2',
                     n.iter = ifelse(test, 100, iter),
                     parallel = TRUE,
                     m.scale = 'log10',
                     num.knots = nk,
                     n.thin = 10,
                     save.sppars = FALSE,
                     plot = FALSE)

plot_indicator(indicator = bma_indicator[,'Index'],
               CIs = bma_indicator[,c(3,4)], 
               main = paste(group, 'indicator', message))

model <- attr(bma_indicator, 'model')
                
if(save_model) 
  if(iter < 20000)
    #saveRDS(object=model, file=file.path('BMA_development/models/',
        # paste0(group,"_",nk,"_",scaling_factor,".rds"))
    save(model, file=file.path(results_folder,
                    paste0(group,"_",nk,"_",scaling_factor, "_",iter/1000,".rData")))
  else write.csv(model$summary, file=file.path(results_folder,
                    paste0(group,"_",nk,"_",scaling_factor, "_",iter/1000,".csv")))

#plot_growthrates(logLambda = model$mean$logLambda,
#                CIs = with(model, 
#                          data.frame(lower=q2.5$logLambda, #upper=q97.5$logLambda)),
#                 main = paste(group,'growth rate', message))

 
# plot the modelled vs observed spindex
#match_data <- m_data
#spmatch <- data.frame(bms=unique(match_data$species), #model=order(unique(match_data$species)))
#match_data$species <- spmatch$model[match(match_data$species, spmatch$bms)]

#plot_species(spindex = model$mean$spindex,
#             lower = model$q2.5$spindex,
#             upper = model$q97.5$spindex,
#             data_in = match_data, main = paste0(group, " indices", message))

############################################################### traceplot

sims.array <- as.data.frame(model$sims.list)
sims.array$chain <- melt(sapply(1:3, rep, times=model$mcmc.info$n.samples/3))$value
sims.array$it <- rep(1:(model$mcmc.info$n.samples/3), times=3)
sims.array <- melt(sims.array, id=c('it', 'chain'))
kit <- floor(model$mcmc.info$n.iter/1000)

gp <- ggplot(data=sims.array, aes(x=it, y=value, col=factor(chain))) +
  facet_wrap(~variable, scales='free', ncol=6) + geom_line() + theme_bw()
gp + ggtitle(paste(group, 'traceplot', message))
ggsave(gp, file=file.path(results_folder, paste('tp',group,nk,scaling_factor,kit,'kit.png', sep="_")),
       he=600, wi=400, units='mm', type='cairo')


###################################################### end loops
}}}
```
